### 2011-11-07 doc/pldi2012/experiments.tex
Some modifications to the experimental section

To measure Legion's ability to distribute work across a distributed machine, 
three benchmarks were developed and run on a small cluster under a variety
of configurations.  The cluster consists of four systems connected by a 40~Gb/s
Infiniband fabric.  Each system is powered by 2 Intel Xeon X5680 CPUs, each
with 6 cores, and 2 NVIDIA Tesla C2070 GPUs.  The systems were running Linux
with GASNet using the {\tt ibv} conduit to interface directly with the standard
OFED software stack for low-latency access to the Infiniband hardware.

@@

The circuit simulator benchmark was designed to be representative of the
general class of finite-element simulations on irregular grids.  It relies
on many iterations using small stencils (e.g. just the neighboring nodes) to
propagate changes through the system, and each iteration requires communication
between neighboring nodes.  Careful partitioning of the system can minimize
the amount of information that must be exchanged, but the irregularity of the
underlying graph mandates similar irregularity in the sharing pattern.

The structure of the circuit simulator was presented in Section~\ref{sec:ex}.
Our implementation uses CUDA to perform the computations on the GPU, and Legion
automatically handles the movement of data between the GPUs' memories and the
system memory, but only when necessary.  In fact, with the help of an
application-specific mapper, it is able to keep much of
the data permanently resident in a given GPU for the entirety of the simulation.
Figure~\ref{sfig:results1:perf} shows the relative speedup as the simulation is
spread over multiple nodes in the cluster.  The size of the circuit
being simulated is not changed as the node count increases.  This is therefore
a demonstration of ``strong scaling''.  Adding a second node to the simulation
speeds it up by 86\%, and with four nodes, the simulation is sped up by a total
of 241\% (i.e. a factor of 3.41x).  When enabling the use of both GPUs in each
node (a decision that requires no changes to the application), the we achieve
a speedup of 5.90x when using all eight GPUs.

@@

Table\~ref{sfig:results1:timers} breaks down the execution time by where it was
spent.  The times are summed over all threads, so the aggregate execution time
will always exceed the elapsed wall clock time.  The time spent in the 
application code is nearly constant in all configurations, but the time spent
copying data jumps as soon as the simulation is spread across multiple nodes.
This is not unexpected, and while the the overhead is responsible for the
less-than-linear performance scaling, the slow growth in the absolute overhead
(after the initial jump) suggests that larger clusters would enjoy additional
speedup, even without increasing the size of the problem being tackled.

@@

To show that Legion remains efficient for regular partitioning cases as well,
we have implemented a particle simulation benchmark based heavily on the
{\tt fluidanimate} benchmark from the PARSEC \cite{bienia11benchmarking} suite.
Our simulation uses a cyclic 2D grid of cells instead of PARSEC's 3D grid with
boundaries, but the computations being performed are preserved.
Although the 
simulation is moving and colliding individual particles, those particles
are placed in cells that make up a regular grid.  The grid is
divided into blocks to be processed by individual threads, with only the
particles in the edge cells being of interest to other threads.  The PARSEC
implementation uses fine-grained locking and the assumption of cache
coherence to access neighboring particles, but neither of these are conducive
to performance scaling across a cluster.  Our fluid benchmark replaces these
accesses with rings of
explicit edge, or \emph{ghost}, cells that are exchanged between adjacent
blocks several times per iteration.  The capturing of these inter-block
interactions through edge cells and the partitioning of those cells into
subregions provides the runtime the ability to schedule data movement in
advance and perform a more precise enforcement of the ordering
constraints necessary for correctness (e.g.
requiring a task to wait only for the subset of tasks in the previous pass 
that computed one of the ghost cells being read by this task).

@@

The second significant jump in application time is when using all 12 CPUs in
each node.  With application tasks running on all the available CPU cores,
there are no ``extra'' cores for running the OS, device drivers, and the like.
For example, GASNet's low-latency implementation counts on a CPU core being
available for 
a thread to spin on message arrival notifications.  Allowing that thread to
sleep (and then wait for the OS rescheduling delay) degrades system performance
much more than the penalty being observed in our fluid simulation.  Now that
one can generally count on having many CPU cores in any new system, it might
make sense to reserve one (or more) in the low-level runtime for background
tasks like these in order to provide more predictable performance on the CPU
cores that are exposed to the high-level runtime.

@@

The circuit simulation was able to count on a good static, if irregular,
partition of its workload, but that is not possible in some cases.  One
such example arises in ray tracing, in which the components of a scene to be
rendered are partitioned in a binary tree to minimize the number of expensive
object intersection tests that must be performed.  By traversing its path 
through the tree, a ray is able to efficiently skip entire chunks of the 
scene that it is guaranteed not to intersect with.  The trees tend to be
highly imbalanced (often intentionally so), and the set of rays that traverse
the tree tend to have similar path.  Both of these factors contribute to
large fluctuations in the load on any particular part of the tree, complicating
attempts to statically partition the tree.  The {\tt Tree Search} benchmark
explores Legion's ability to load-balance tasks in a locality-aware manner.

>>>>

To measure Legion's ability to distribute work across a distributed machine, 
three benchmarks were developed and run on a small cluster under a variety
of configurations.  The cluster consists of four nodes connected by a 40~Gb/s
Infiniband fabric.  Each node is powered by 2 Intel Xeon X5680 CPUs, each
with 6 cores, and 2 NVIDIA Tesla C2070 GPUs.  The cluster is running Linux
with GASNet using the {\tt ibv} conduit to interface directly with the standard
OFED software stack for low-latency access to the Infiniband hardware.

@@

The circuit simulator benchmark was designed to be representative of the
general class of finite-element simulations on irregular grids.  It relies
on many iterations using small stencils (e.g. just the neighboring nodes) to
propagate changes through the system.  Each iteration requires communication
between neighboring nodes.  Careful partitioning of the system can minimize
the amount of information that must be exchanged, but the irregularity of the
underlying graph mandates similar irregularity in the sharing pattern.

The structure of the circuit simulator was presented in Section~\ref{sec:ex}.
Our implementation uses CUDA to perform the computations on the GPU, and Legion
automatically handles the movement of data between the GPUs' memories and the
system memory when communication is necessary.  By utilizing an  
application-specific mapper, our implementation is able to keep much of
the data permanently resident in a given GPU for the entirety of the simulation.
Figure ~\ref{sfig:results1:perf} shows the relative speedup as the simulation is
spread over multiple nodes in the cluster.  We demonstrate strong scaling by
keeping the size of the circuit fixed as node count increases.
Adding a second node to the simulation
speeds it up by 86\%, and with four nodes, the simulation is sped up by a total
of 241\% (i.e. a factor of 3.41x).  When enabling the use of both GPUs in each
node (a decision that requires no changes to the application), we achieve
a speedup of 5.90x when using all eight GPUs.

@@

Table~\ref{sfig:results1:timers} breaks down the execution time by where it was
spent.  The times are summed over all threads, so the aggregate execution time
will always exceed the elapsed wall clock time.  The time spent in the 
application code is nearly constant in all configurations, but the time spent
copying data jumps as soon as the simulation is spread across multiple nodes.
This is not unexpected, and while the the overhead is responsible for the
less-than-linear performance scaling, the slow growth in the absolute overhead
(after the initial jump) suggests that larger clusters would enjoy additional
speedup, even without increasing the size of the problem being tackled.

@@

To show that Legion remains efficient for regular partitioning cases as well,
we have implemented a particle simulation benchmark based on the
{\tt fluidanimate} benchmark from the PARSEC \cite{bienia11benchmarking} suite.
Our simulation uses a cyclic 2D grid of cells instead of PARSEC's 3D grid with
boundaries, but the computations being performed are preserved.
Although the 
simulation is moving and colliding individual particles, those particles
are placed in cells that make up a regular grid.  The grid is
divided into blocks to be processed by individual threads, with only the
particles in the edge cells being of interest to other threads.  The PARSEC
implementation uses fine-grained locking and the assumption of cache
coherence to access neighboring particles, but neither of these are conducive
to performance scaling across a cluster.  Our fluid benchmark replaces these
accesses with rings of
explicit edge, or \emph{ghost}, cells that are exchanged between adjacent
blocks several times per iteration.  The capturing of these inter-block
interactions through edge cells and the partitioning of those cells into
subregions provides the runtime the ability to schedule data movement in
advance and perform a more precise enforcement of the ordering
constraints necessary for correctness (e.g.
requiring a task to wait only for the subset of tasks in the previous pass 
that computed one of the ghost cells being read by this task).

@@

The second significant jump in application time is when using all 12 CPUs in
each node.  With application tasks running on all the available CPU cores,
there are no extra cores for running the OS or device drivers.
For example, GASNet's low-latency implementation counts on a CPU core being
available for 
a thread to spin on active message arrival notifications.  Allowing that thread to
sleep (and then wait for the OS rescheduling delay) degrades system performance
much more than the penalty being observed in our fluid simulation.  In future
work we plan to reserve low-level runtime processors for these background
tasks to provide more predictable performance when using all CPU cores.

%Now that
%one can generally count on having many CPU cores in any new system, it might
%make sense to reserve one (or more) in the low-level runtime for background
%tasks like these in order to provide more predictable performance on the CPU
%cores that are exposed to the high-level runtime.

@@

The circuit simulation was able to count on a static, if irregular,
partition of its workload, but that is not possible in some cases.  One
such example arises in ray tracing, in which the components of a scene to be
rendered are partitioned in a binary tree to minimize the number of expensive
object intersection tests that must be performed.  By traversing its path 
through the tree, a ray is able to efficiently skip entire chunks of the 
scene that it is guaranteed not to intersect with.  The trees tend to be
highly imbalanced (often intentionally so), and the set of rays that traverse
the tree tend to have similar path.  Both of these factors contribute to
large fluctuations in the load on any particular part of the tree, complicating
attempts to statically partition the tree.  The {\tt Tree Search} benchmark
explores Legion's ability to load-balance tasks in a locality-aware manner.
